%! TEX program = xelatex
\documentclass{article}
%\documentclass[acmsmall,screen,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\input{preamble.tex}

\title{Modular Analysis}
\author{Joonhyup Lee}
\begin{document}
\maketitle
\begin{abstract}
  We present a framework of modular static analysis that analyzes open
  program fragments in advance and complete the whole analysis later
  when the fragments are closed. The framework is defined for a
  call-by-value lambda calculus extended with constructs for defining and
  linking first-class modules (a collection of bindings) that support
  recursive bindings for values, modules, and functors(module
  functions). Thanks to the abstract interpretation framework, our
  modular analysis framework's key is how to define the semantics,
  called ``shadow'' semantics of open program fragments so that the
  computation with free variables should be captured in advance
  and be completed later at link-time. A modular
  static analysis is abstractions of this shadow semantics and of the
  linking operation. The safety of the framework is proven in Coq.
  Two instances of the framework are presented: value analyses for an
  applicative language(control-flow, closure analysis) and for
  an imperative language.
  %  The framework is intended to be a guide in
  %  designing modular analyses in practice for the specific source language at hand. 
\end{abstract}

\section{Problem}
By modular analysis we mean a static analysis technology that analyzes
programs in the compositional way. Modular analysis partially analyzes
program fragments separately and then complete the analysis when all
the fragments are available and linked.

Designing a modular analysis decides on two factors: how to build
sound partial semantic summaries of program fragments and how to
complete them at link time. A particular modular analysis strikes a
balance between the two factors. Depending on how much analysis work
is invested on each factor, modular analyses range from one extreme, a
trivial one (wait until the whole code is available and do the
whole-program analysis) to another extreme, non-trivial yet
hardly-automatic ones (abduct assumptions about free variables and
check them at link-time).

However, designing a modular analysis has been elusive. There is a
lack of a general framework by which we can design sound modular
analyses with varying balance and accuracy of our choice.
We need a general framework by which static analysis designers can
design non-trivial sound modular analyses.

\section{Framework Sketch}
Thanks to the abstract interpretation framework, the key in our
modular analysis framework is how to define the semantics (we call it
``shadow'' semantics) of open program fragments so that the
computation with free variables should be captured in advance
and be completed later at link-time. A modular static analysis is
then nothing but abstractions of this shadow semantics and of the
linking operation.

\subsection{Examples of Shadow Semantics and its Abstraction}
Shadow semantics is the semantics for the computation involving
free variables. We give examples of shadows for program fragments
written in OCaml.

\paragraph{Example 1: Mapping an Unknown Function Over a Known List}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let rec map = fun f -> fun l ->
  match \textsuperscript{0}l with
  | [] -> []
  | hd :: tl -> \textsuperscript{1}(\textsuperscript{3}f \textsuperscript{4}hd) :: \textsuperscript{2}(map f tl)
in map \textsuperscript{5}g \textsuperscript{6}(1 :: 2 :: 3 :: [])
  \end{BVerbatim}
\end{center}
The shadow semantics of the above fragment is
\[
  \{\CallE(\abbrtext{G},1)\cons\CallE(\abbrtext{G},2)\cons\CallE(\abbrtext{G},3)\cons[]\}
  \text{ where }
  \abbrtext{G}=\ReadE(\InitE,\texttt{g})
\]
Shadows such as $\CallE$, $\ReadE$ correspond to semantic operations,
like function application and reading from the environment.
The $\InitE$ shadow is the unknown initial environment that the fragment will execute under.

Computing sound and finite approximations of the shadow
semantics of program fragments corresponds to building sound partial
semantic summaries in advance. For the above example,
we may use an abstraction that stores summaries of each input \emph{environment}
and output \emph{value} per program point.

\vspace{\topsep}
\noindent\textbf{Notation.}
We write $\ppoint_i$ for each program point $i$.
$\ppoint_i.\abbrtext{in}$ denotes the input to $\ppoint_i$, and $\ppoint_i.\abbrtext{out}$
denotes the output from $\ppoint_i$.
For brevity, we omit the selector in $\ppoint_i.\abbrtext{out}$ and write $\ppoint_i$.
We write $[a,b]_{\abbrtext{itv}}$ to denote the interval from integer $a$ to $b$.

\vspace{\topsep}
Now, the abstract shadow
\[\{\ppoint_1\cons \ppoint_2\}\]
is a sound approximation of the concrete shadow. To illustrate why,
here are abstract shadows from some program points involved in the map computation.
\begin{align*}
  \ppoint_1.\abbrtext{out}:\{ & \Abs\CallE(\ppoint_3,\ppoint_4)\}                &  &  & \ppoint_4.\abbrtext{out}: & \;[1,3]_{\abbrtext{itv}}                                                             \\
  \ppoint_2.\abbrtext{out}:\{ & [],\ppoint_1\cons \ppoint_2\}                    &  &  & \ppoint_5.\abbrtext{in}:  & \;\{\Abs\InitE,\texttt{map}\mapsto\{\langle\lambda\texttt{f}.\texttt{...}\rangle\}\} \\
  \ppoint_3.\abbrtext{out}:\{ & \Abs\ReadE(\ppoint_5.\abbrtext{in},\texttt{g})\}
\end{align*}
Starting from $\ppoint_1\cons\ppoint_2$, we can reach the concrete shadow by
expanding each program point according to the above table.

\paragraph{Example 2: Mapping an Unknown Function Over an Unknown List}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let rec map = ... in map \textsuperscript{5}g \textsuperscript{6}lst
  \end{BVerbatim}
\end{center}
A more interesting case is when even the list is replaced with a free variable.
The shadow semantics of the above fragment is
\begin{center}
  \begin{tabular}{c}
    $\{[],\CallE(\abbrtext{G},\Hd(\abbrtext{Lst}))\cons[],\CallE(\abbrtext{G},\Hd(\abbrtext{Lst}))\cons\CallE(\abbrtext{G},\Hd(\Tl(\text{Lst})))\cons[],$ \\
    [4pt]
    $\dots,\CallE(\abbrtext{G},\Hd(\abbrtext{Lst}))\cons\cdots\cons\CallE(\abbrtext{G},\Hd(\Tl^{n}(\abbrtext{Lst})))\cons[],\dots\}$
  \end{tabular}
\end{center}
where
\begin{align*}
  \abbrtext{G} & =\ReadE(\InitE,\texttt{g})       &  &  & \abbrtext{Lst} & =\ReadE(\InitE,\texttt{lst})     \\
  \Hd(l)       & =\textsf{Match}(l,\texttt{::},0) &  &  & \Tl(l)         & =\textsf{Match}(l,\texttt{::},1)
\end{align*}

The same abstraction as in Example 1 results in the same abstract shadow
\[\{\ppoint_1\cons\ppoint_2\}\]
but with different shadows for each program point.
\begin{align*}
  \ppoint_0.\abbrtext{out} :\{ & \Abs\ReadE(\ppoint_6.\abbrtext{in},\texttt{lst}), &  &  & \ppoint_3.\abbrtext{out}:\{ & \Abs\ReadE(\ppoint_5.\abbrtext{in},\texttt{g})\}                                 \\
                               & \Abs{\textsf{Match}}(\ppoint_0,\texttt{::},1)\}   &  &  & \ppoint_4.\abbrtext{out}:\{ & \Abs{\textsf{Match}}(\ppoint_0,\texttt{::},0)\}                                  \\
  \ppoint_1.\abbrtext{out} :\{ & \Abs\CallE(\ppoint_3,\ppoint_4)\}                 &  &  & \ppoint_5.\abbrtext{in} :\{ & \Abs\InitE,\texttt{map}\mapsto\{\langle\lambda\texttt{f}.\texttt{...}\rangle\}\} \\
  \ppoint_2.\abbrtext{out} :\{ & [],\ppoint_1\cons \ppoint_2\}                     &  &  & \ppoint_6.\abbrtext{in} :\{ & \Abs\InitE,\texttt{map}\mapsto\{\langle\lambda\texttt{f}.\texttt{...}\rangle\}\}
\end{align*}

\paragraph{Example 3: Unknown Function Modifying the Memory}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let x = \textsuperscript{1}ref 1 in \textsuperscript{2}(\textsuperscript{4}g \textsuperscript{5}x); \textsuperscript{3}!x
  \end{BVerbatim}
\end{center}

The shadow semantics can be extended to support imperative features.
\[\texttt{ref }e\text{ (allocation)}\quad x\texttt{ := }e\text{ (update)}\quad\texttt{!}x\text{ (dereference)}\]
We only have to add a memory component, and a shadow for \emph{dereferencing}.

In the above program, the environment and memory at $\ppoint_1$, $\ppoint_2$, and $\ppoint_3$ are
\[
  \setlength{\arraycolsep}{2pt}
  \begin{array}{c|c|c}
              & \text{Environment}                                     & \text{Memory}                                                                                                       \\
    \hline
    \ppoint_1 & \langle\varnothing,\InitE\rangle                       & \langle\varnothing,\InitE\rangle                                                                                    \\
    \ppoint_2 & \langle\{\texttt{x}\mapsto\abbrtext{a}\},\InitE\rangle & \langle\{\abbrtext{a}\mapsto 1\},\InitE\rangle\text{ where }\abbrtext{a}\in\Addr                                    \\
    \ppoint_3 & \langle\{\texttt{x}\mapsto\abbrtext{a}\},\InitE\rangle & \langle\{\abbrtext{a}\mapsto \DerefAE(C,\abbrtext{a})\},C\rangle\text{ where }C=\CallE(\abbrtext{G},\abbrtext{a}|m)
  \end{array}
\]
where
\[\abbrtext{G}=\ReadE(\InitE,\texttt{g})\text{ and }m=\langle\{\abbrtext{a}\mapsto 1\},\InitE\rangle\]
The notation $\CallE(\abbrtext{G},\abbrtext{a}|m)$ is inspired from conditional probability,
and means that the shadow $\abbrtext{G}$ is called with argument $\abbrtext{a}$ \emph{given}
memory $m$.

The memory is a pair of (1) the set of bindings from addresses to values,
and (2) the shadow that keeps track of unknown modifications to the memory.
Here, the unknown function call $\CallE(\abbrtext{G},\abbrtext{a}|m)$
modifies what was stored in $\abbrtext{a}$.

\vspace{\topsep}
\noindent\textbf{Notation.}
We write $\Abs{\abbrtext{a}}_i$ to abstract the concrete location
allocated at $\ppoint_i$.
$\ppoint_i.\abbrtext{in}.\Abs\ctx$ denotes the input environment
at $\ppoint_i$,
and $\ppoint_i.\abbrtext{in}.\Abs\mem$ denotes the input memory.
Likewise, $\ppoint_i.\abbrtext{out}.\Abs{v}$ denotes
the output value at $\ppoint_i$,
and $\ppoint_i.\abbrtext{out}.\Abs\mem$ denotes
the output memory.
The selectors $.\abbrtext{out}$, $.\Abs\ctx$, $.\Abs{v}$, and $.\Abs\mem$
are omitted.

\vspace{\topsep}
Using the same na\"ive abstraction as before, the abstract result is
\[\{\Abs{\DerefAE}(\ppoint_2,\ppoint_1)\}\quad(\text{abbreviated form of }\Abs\DerefAE(\ppoint_2.\abbrtext{out}.\Abs{m},\ppoint_1.\abbrtext{out}.\Abs{v}))\]
where
\begin{align*}
  \ppoint_2.\abbrtext{out}.\Abs{v} :\{  & \Abs\CallE(\ppoint_4,\ppoint_5|\ppoint_5)\}                        &  &  & \ppoint_4.\abbrtext{out}.\Abs{v}:\{  & \Abs\ReadE(\ppoint_4.\abbrtext{in},\texttt{g})\}                      \\
  \ppoint_2.\abbrtext{out}.\Abs\mem :\{ & \Abs\CallE(\ppoint_4,\ppoint_5|\ppoint_5),                         &  &  & \ppoint_5.\abbrtext{in}.\Abs\ctx :\{ & \Abs\InitE,\texttt{x}\mapsto\{\Abs{\abbrtext{a}}_1\}\}                \\
                                        & \Abs{\abbrtext{a}}_1\mapsto\{\Abs\DerefAE(\ppoint_2,\ppoint_1)\}\} &  &  & \ppoint_5.\abbrtext{out}.\Abs{v}:\{  & \Abs{\abbrtext{a}}_1,\Abs\ReadE(\ppoint_5.\abbrtext{in},\texttt{x})\} \\
  \ppoint_4.\abbrtext{in}.\Abs\ctx :\{  & \Abs\InitE,\texttt{x}\mapsto\{\Abs{\abbrtext{a}}_1\}\}             &  &  & \ppoint_5.\abbrtext{out}.\Abs\mem:\{ & \Abs\InitE,\Abs{\abbrtext{a}}_1\mapsto[1,1]_{\abbrtext{itv}}\}
\end{align*}

\paragraph{Example 4: Reading and Writing to Shadow Addresses}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let t = (!x)\textsuperscript{1} in (x := (!y)\textsuperscript{2})\textsuperscript{3}; (y := t)\textsuperscript{4}
  \end{BVerbatim}
\end{center}

The addresses that are dereferenced and updated may also be shadows.
Thus, the memory must store bindings from \emph{shadow addresses} to values.
A shadow address is a \emph{set} consisting of one or less actual address and zero
or more shadows that are aliases of the actual address.

Upon exit at each program point marked above,
\[
  \setlength{\arraycolsep}{2pt}
  \begin{array}{c|c|c}
              & \text{Environment}                                      & \text{Possible Memories}                                                                                                                                                                \\
    \hline
    \ppoint_1 & \langle\{\texttt{t}\mapsto*\abbrtext{X}\},\InitE\rangle & \langle\{\{\abbrtext{X}\}\mapsto*\abbrtext{X}\},\InitE\rangle                                                                                                                           \\
    \ppoint_2 & \langle\{\texttt{t}\mapsto*\abbrtext{X}\},\InitE\rangle & \langle\{\{\abbrtext{X},\abbrtext{Y}\}\mapsto*\abbrtext{X}\},\InitE\rangle\text{ or }\langle\{\{\abbrtext{X}\}\mapsto*\abbrtext{X},\{\abbrtext{Y}\}\mapsto*\abbrtext{Y}\},\InitE\rangle \\
    \ppoint_3 & \langle\{\texttt{t}\mapsto*\abbrtext{X}\},\InitE\rangle & \langle\{\{\abbrtext{X},\abbrtext{Y}\}\mapsto*\abbrtext{X}\},\InitE\rangle\text{ or }\langle\{\{\abbrtext{X}\}\mapsto*\abbrtext{Y},\{\abbrtext{Y}\}\mapsto*\abbrtext{Y}\},\InitE\rangle \\
    \ppoint_4 & \langle\{\texttt{t}\mapsto*\abbrtext{X}\},\InitE\rangle & \langle\{\{\abbrtext{X},\abbrtext{Y}\}\mapsto*\abbrtext{X}\},\InitE\rangle\text{ or }\langle\{\{\abbrtext{X}\}\mapsto*\abbrtext{Y},\{\abbrtext{Y}\}\mapsto*\abbrtext{X}\},\InitE\rangle
  \end{array}
\]
where
\begin{align*}
  \abbrtext{X}=\ReadE(\InitE,\texttt{x})\quad\abbrtext{Y}=\ReadE(\InitE,\texttt{y})\quad*\!S=\DerefSE(\InitE,S)\text{ for shadow }S
\end{align*}

Unlike the previous example, there are multiple memories possible after
execution. Each memory represents different aliasing assumptions.
For example, $\langle\{\{\abbrtext{X},\abbrtext{Y}\}\mapsto\_\},\_\rangle$
encodes the assumption that \texttt{x} and \texttt{y} are aliases.

Computing an abstraction of the shadow memory is an interesting
problem. One must retain some information about the aliasing
\emph{invariant} that is maintained throughout the execution of the
program to avoid degenerating to a meaningless analysis.

One traditional way is by assuming a partition of variables that
distinguishes variables that must-not-alias.
For the above program, we may assume the invariant that \texttt{x} and
\texttt{y} must not alias.

\vspace{\topsep}
\noindent\textbf{Notation.}
The memory is abstracted by a set of abstract shadows and
abstract bindings of the form $*x\mapsto(\text{set of abstract values})$.
The abstract shadow $\Abs\DerefAE(\ppoint_i,x)$ concretizes to
$\DerefAE(S,\ctx(x))$, where $S$ is abstracted by $\ppoint_i.\abbrtext{in}.\Abs\mem$
and $\ctx$ is abstracted by $\ppoint_i.\abbrtext{in}.\Abs\ctx$.

\vspace{\topsep}
Assuming a flow-sensitive analysis
that allows strong updates when the updated variable can have only
one concrete address,
\[
  \setlength{\arraycolsep}{2pt}
  \begin{array}{c|c|c}
                             & .\Abs\ctx                                               & .\Abs\mem                                                                                          \\
    \hline
    \ppoint_2.\abbrtext{in}  & \{\texttt{t}\mapsto\{*\Abs{\abbrtext{X}}\},\Abs\InitE\} & \{*\texttt{x}\mapsto\{*\Abs{\abbrtext{X}}\},\Abs\InitE\}                                           \\
    \ppoint_3.\abbrtext{in}  & \{\texttt{t}\mapsto\{*\Abs{\abbrtext{X}}\},\Abs\InitE\} & \{*\texttt{x}\mapsto\{*\Abs{\abbrtext{X}}\},*\texttt{y}\mapsto\{*\Abs{\abbrtext{Y}}\},\Abs\InitE\} \\
    \ppoint_4.\abbrtext{in}  & \{\texttt{t}\mapsto\{*\Abs{\abbrtext{X}}\},\Abs\InitE\} & \{*\texttt{x}\mapsto\{*\Abs{\abbrtext{Y}}\},*\texttt{y}\mapsto\{*\Abs{\abbrtext{Y}}\},\Abs\InitE\} \\
    \ppoint_4.\abbrtext{out} & -                                                       & \{*\texttt{x}\mapsto\{*\Abs{\abbrtext{Y}}\},*\texttt{y}\mapsto\{*\Abs{\abbrtext{X}}\},\Abs\InitE\}
  \end{array}
\]
where
\[*\Abs{\abbrtext{X}}=\Abs\DerefAE(\ppoint_1,\texttt{x})\quad *\Abs{\abbrtext{Y}}=\Abs\DerefAE(\ppoint_2,\texttt{y})\]
The aliasing assumption is used at $\ppoint_2$, where dereferencing \texttt{y}
doesn't include the value pointed by $\texttt{x}$.
Strong updates are possible, since \texttt{x} concretizes only to $\ReadE(\InitE,\texttt{x})$,
and \texttt{y} concretizes only to $\ReadE(\InitE,\texttt{y})$ throughout the analysis.

\subsection{Examples of Linking and its Abstraction}
The shadow semantics become actual when the involved free variables
are known at link time. The linking semantics, the semantics of the link
operation, defines this actualization operation. For the above fragments, let's
consider that a closing fragment is available.

\paragraph{Example 5: Module Exporting a User-Defined Function}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let g = \textsuperscript{7}fun x -> x + 1
  \end{BVerbatim}
\end{center}
This is a module which returns the environment
$\ctx=\{\texttt{g}\mapsto\langle\lambda\texttt{x}.\texttt{x + 1},\varnothing\rangle\}$.
Linking this with the concrete shadow in Example 1 gives:
\[\{2\cons3\cons4\cons[]\}\]

Computing a sound and finite approximation of the linking semantics
corresponds to completing partial analysis summaries at link time when
the analysis results for the involved free variable are available. For
the above example,
\[
  \Abs\ctx=\{\texttt{g}\mapsto\{\langle\lambda\texttt{x}.\texttt{x + 1},\ppoint_7.\abbrtext{in}\rangle\}\}\text{ where }
  \ppoint_7.\abbrtext{in}:\varnothing
\]
is the abstract shadow.

Applying a sound abstract version of the linking operator will result in
\begin{align*}
  \ppoint_1.\abbrtext{out} & :[2,4]_{\abbrtext{itv}}                                                     &  &  & \ppoint_4.\abbrtext{out} & :[1,3]_{\abbrtext{itv}} \\
  \ppoint_2.\abbrtext{out} & :\{[],\ppoint_1\cons \ppoint_2\}                                            &  &  & \ppoint_5.\abbrtext{in}  & :\Abs\ctx               \\
  \ppoint_3.\abbrtext{out} & :\{\langle\lambda\texttt{x}.\texttt{x + 1},\ppoint_7.\abbrtext{in}\rangle\} &  &  &
\end{align*}

\paragraph{Example 6: Module Exporting a Foreign Function}
\begin{center}
  \begin{BVerbatim}
external g : int -> int = "incr"
  \end{BVerbatim}
\end{center}
The function \texttt{g} might also be a foreign function.
The return value of this module is
$\ctx=\{\texttt{g}\mapsto\textsf{Prim}(\texttt{incr})\}$,
where \textsf{Prim} stands for a \emph{primitive} value.
Linking this with the concrete shadow in Example 1 gives:
\[\{\textsf{PrimCall}(\texttt{incr},1)\cons\textsf{PrimCall}(\texttt{incr},2)\cons\textsf{PrimCall}(\texttt{incr},3)\cons[]\}\]

The sound abstraction for $\ctx$ is
\[
  \Abs\ctx=\{\texttt{g}\mapsto\{\textsf{Prim}(\texttt{incr})\}\}
\]
Applying a sound abstract version of the linking operator will result in
\begin{align*}
  \ppoint_1.\abbrtext{out} & :\{\Abs{\textsf{PrimCall}}(\texttt{incr},\ppoint_4)\} &  &  & \ppoint_4.\abbrtext{out} & :[1,3]_{\abbrtext{itv}} \\
  \ppoint_2.\abbrtext{out} & :\{[],\ppoint_1\cons \ppoint_2\}                      &  &  & \ppoint_5.\abbrtext{in}  & :\Abs\ctx               \\
  \ppoint_3.\abbrtext{out} & :\{\textsf{Prim}(\texttt{incr})\}
\end{align*}

\paragraph{Example 7: Partial Resolution of Shadows}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let lst = \textsuperscript{7}1 :: \textsuperscript{8}(\textsuperscript{9}2 :: \textsuperscript{10}(\textsuperscript{11}3 :: \textsuperscript{12}[]))
  \end{BVerbatim}
\end{center}
The return value of this module is
$\ctx=\langle\{\texttt{lst}\mapsto 1\cons2\cons3\cons[]\},\InitE\rangle$.
Linking this with the concrete shadow in Example 2 gives:
\[
  \{[],\CallE(\abbrtext{G},1)\cons[],\CallE(\abbrtext{G},1)\cons\CallE(\abbrtext{G},2)\cons[],
  \CallE(\abbrtext{G},1)\cons\CallE(\abbrtext{G},2)\cons\CallE(\abbrtext{G},3)\cons[]\}
\]
where
\[
  \abbrtext{G}=\ReadE(\InitE,\texttt{g})
\]

The reason for this imprecision is because some information is forgotten.
For example, the output of Example 2 can be $[]$ if and only if
$\abbrtext{Lst}$ is matched with $[]$. To prevent this, shadows may be augmented with
constraints, such as $\textsf{Matched}(\abbrtext{Lst},[])$, but we elide this detail
for presentation.

The sound abstraction for $\ctx$ is
\[
  \Abs\ctx=\{\Abs\InitE,\texttt{lst}\mapsto\{\ppoint_7\cons\ppoint_8\}\}
  \text{ where }
  \ppoint_7.\abbrtext{out}:[1,1]_{\abbrtext{itv}}
  \text{, }
  \ppoint_8.\abbrtext{out}:\{\ppoint_9\cons\ppoint_{10}\}\dots
\]
The result of linking $\Abs\ctx$ to the abstract shadow in Example 2
is equivalent to the result of Example 1. This means that the
shadows can be filled in \emph{incrementally}.

\paragraph{Example 8: Linking into Memory}
\begin{center}
  \begin{BVerbatim}[commandchars=\\\{\}]
let x = \textsuperscript{5}ref 1
let y = \textsuperscript{6}ref 2
  \end{BVerbatim}
\end{center}

Assume we want to link this module with Example 4.
The environment and memory exported by this module are
\[\ctx=\{\texttt{x}\mapsto a_5,\texttt{y}\mapsto a_6\}\quad\mem=\{\{a_5\}\mapsto 1,\{a_6\}\mapsto 2\}\quad\text{where }a_5\neq a_6\]
Linking is extended, where linking an input environment and memory
with a shadow results in an output value and memory.

Since the result of linking into $\abbrtext{X}=\ReadE(\InitE,\texttt{x})$ and
$\abbrtext{Y}=\ReadE(\InitE,\texttt{y})$ are different,
only the memories where $\abbrtext{X}$ and $\abbrtext{Y}$ are kept
separate are resolved.
The result of concrete linking is thus:
\[
  \setlength{\arraycolsep}{2pt}
  \begin{array}{c|c|c}
              & \text{Environment}      & \text{Possible Memories}              \\
    \hline
    \ppoint_4 & \{\texttt{t}\mapsto 1\} & \{\{a_5\}\mapsto 2,\{a_6\}\mapsto 1\}
  \end{array}
\]

To link with the abstract memory in Example 4, the module
must be analyzed under the same assumptions.
Namely, \texttt{x} and \texttt{y} should be always assumed
to be not aliasing.
Under this assumption,
\[
  \Abs\ctx=\{\texttt{x}\mapsto\{\Abs{a}_5\},\texttt{y}\mapsto\{\Abs{a}_6\}\}\quad
  \Abs\mem=\{*\texttt{x}\mapsto[1,1]_{\abbrtext{itv}},*\texttt{y}\mapsto[2,2]_{\abbrtext{itv}}\}
\]
and linking results in the expected behavior.

\subsection{Paper Overview}
We present our framework for a call-by-value lambda calculus extended
with constructs for defining and linking first-class modules (a
collection of bindings) that support recursive bindings for values,
modules, and functors(module functions).

The framework shows two points: how to define the shadow semantics and
what to prove for the soundness of consequent modular analysis. The
safety of the framework is proven in Coq. We present two instances of the
framework: for high-order applicative language we show modular
closure analysis design, and for imperative languages we show modular
  [TODO] analysis design.

\clearpage
\input{semantics.tex}
\clearpage
\input{linking.tex}
\clearpage
\input{cfa.tex}

\section{Conclusion}

\end{document}

